Sentiment classification of online political discussion -- a comparison of two methods

### 2014-03-20 manuscript.tex
Corrected a typo.
keyword to target + minor changes

keywords: \textit{}

@@

In this paper we focus on classifying the sentiment towards
religious/political topics, say the Quran, in Norwegian political
discussion. We use a lexicon-based approach where we classify the
sentiment of a sentence based on the polarity of sentiment words in
relation to a set of target words. We expect that statistically the
importance of a sentiment word towards the keyword is related to the
number of words between the sentiment and key word as suggested by
Ding et al. \shortcite{Ding08}.  Information about the syntactic
environment of certain words or phrases has in previous work also been
shown to be useful for the task of sentiment classification
\cite{Wil:Wie:Hof:09,Jiang11}. In this work we therefore compare the
results obtained using a token-based distance measure with a novel
syntax-based distance measure obtained using dependency graphs and
further augmented with linguistically motivated syntactic patterns
expressed as dependency paths.  We furthermore present a freely
available corpus of Norwegian political discussion related to religion
and immigration, which has been manually annotated for the sentiment
expressed towards a set of target words, as well as a manually translated sentiment lexicon.

%% In this paper we focus on classifying the opinion toward religious/political topics, say the Quran, in political discussion by using the lexical-based approach. One intuitive approach is to find both the keyword (e.g. Quran) and the words with sentiment in the sentence and classify the sentiment of the sentence based on the polarity of these sentiment words. We expect that statistically the importance of a sentiment word towards the keyword is related to the number of words between the sentiment and key word as suggested by Ding et al. \shortcite{Ding08}. Two other approaches is to automatically parse the material and either use the distance between key and sentiment word in the parse tree or develop grammatical dependence paths, see e.g. Jiang et al. \shortcite{Jiang11}. The aim of this paper is to compare the performance of a word distance method \cite{Ding08} with novel methods based on distance in parse tree and grammatical dependence paths to classify opinions in political discussions.

@@

In this Section we present two methods to classify sentences as either positive, neutral or negative towards a keyword. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word (sentiment lexicon). Both keywords, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters are words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. Let $kw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the keyword in the sentence. Further let $sw_{j}, j \in \{1,2,\ldots,J\}$ be appearance number $j$ of a sentiment word in the sentence. Finally let $\textbf{ss} = (ss_1, ss_2, \ldots, ss_K)$ repressent the sentiment shifters in the sentence. We compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \mathbf{imp}(kw_i, sw_{j})\mathbf{shift}(sw_{j}, \mathbf{ss})    
\end{equation}
where the function \textbf{imp} computes the importance of the sentiment word $sw_{j}$ on the keyword appearance $kw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{shift}(sw_{j}, \mathbf{ss})$ computes whether the sentiment of $sw_{j}$ should be shifted based on all the sentiment shifters in the sentence. It returns $-1$ (sentiment shift) if some of the sentiment shifters is within $d_{p}$ words in front or $d_{n}$ words behind $sw_{j}$, respectively. Else the function, returns $1$ (no sentiment shift). We classify the sentiment towards the keyword to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set.

@@

\paragraph{Dependency distance} One way of expressing the syntactic environment of a target word with respect to a sentiment word is to determine its distance in the dependency graph. We therefore define a distance function $\mathbf{depdist}(kw_i, sw_{j})$ which returns the number of nodes in the shortest dependency path from the target word to the sentiment word in the dependency. The shortest path is determined using Dijkstra's shortest path algorithm \cite{Dij:59}.

\paragraph{Dependency paths}
A second way of determining the importance of a sentiment word towards a target based on syntactically parsed texts, is to establish a list of grammatical dependency paths between words, and test whether such paths exist between the targets and sentiment words \cite{Jiang11}. The assumption would be that two words most likely are semantically related to each other if there is a meaningful grammatical relation between them. Furthermore, it is reasonable to expect than some paths are stronger indicators of the overall sentiment of the sentence than others. To test this method, we have made a list of 42 grammatical dependency paths, devided into four groups, and gave them a score between 1-3. The higher the score is, the better indicator of sentiment the path is assumed to be. In the following paragraphs, we will briefly present the groups of paths and the maximum score we have assigned in each group. The paths are written as follows: postag-target:postag-sentiment word{\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn({\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn etc.). \emph{Up} and \emph{dn} show whether you move up or down in the dependency tree.

@@

A second group we have considered, contains paths from subject targets to sentiment words embedded within the predicate, such as from the subject to the nominal direct object of a verb, subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn. Paths from subjects into different kinds of adverbials are also a part of this group. We consider paths from subjects to objects good indicators of sentiment and give them the highest score, 3, based on the same arguments as for subject-predicate paths. Paths into averbials get lower values, as adverbials often are less semantically connected to the predicate than objects.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline \bf Path group & \bf Number & \bf Score range \\ \hline
Subj. to pred. & 9 & 3 \\
Subj. to pred.-internal & 13 & 3-1 \\
Pred.-internal & 6 & 1 \\
NP-internal & 14 & 2-1 \\
\hline
\end{tabular}
\end{center}
\caption{\label{deppath} Dependency paths}
\end{table}

\paragraph{Modified scoring function}
Let $\mathcal{D}$ denote the set of all important dependency paths. The function $\mathbf{gram}(kw_i, sw_{j})$ returns the dependency path, and if $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(kw_i, sw_{j}) \in [0,1]$, returns the importance of the path. Further let $\mathbf{depdist}(kw_i, sw_{j})$ return the dependency distance, as described above. The \textbf{imp} function is computed as follows. If $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$ we use
\begin{align}
  \begin{split}
    \label{eq:3}
  \mathbf{imp}&(kw_i, sw_{j}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{j}) W_{\mathbf{dep}}(kw_i, sw_{j}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is a parameter that weights the score from the important dependency path and the tree distance and can be tuned using a training set. If $\mathbf{gram}(kw_i, sw_{j}) \not\in \mathcal{D}$ we simply use 
\begin{equation}
  \label{eq:4}
  \mathbf{imp}(kw_i, sw_{j}) = \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{treedist}(kw_i, sw_{j})}    
\end{equation}
Note that when $\alpha = 0$, \eqref{eq:3} reduces to \eqref{eq:4}.

@@

We did not find any suitable annotated text material related to political discussions and therefore created our own. We manually selected 46 debate articles from the Norwegian online newspapers \textit{NRK Ytring, Dagbladet, Aftenposten, VG and Bergens Tidene}. To each debate article there were attached a discussion thread where readers could express their opinions and feelings towards the content of the debate article. All the text from the debate articles and the subsequent discussions were collected using text scraping \cite{Hammer13}. The debate articles were related to religion and immigration and we wanted to classify the sentiment towards all words with stem: \textit{islam, muslim, quran, allah, muhammed, imam and mosque}. These are topics that typically create a lot of active discussions and disagreements.

We automatically divided the material into sentences and all sentences containing at least one keyword and one sentiment word were kept for further analysis. If a sentence contained more than one keyword, e.g. both Islam and Quran, the sentence were repeated one time for each keyword in the final text material. We could then classify the sentiment towards each of the keywords in the sentence consecutively. To assure that we do not underestimate the uncertainty in the statistical analysis, we see each repetition of the sentence as the same sentence with respect to the sentence random effect in the regression model in Section \ref{sec:sa}

Each sentence was manually annotated as to whether the commenter was positive, negative or neutral towards the keyword in the sentence. Each sentence was evaluated individually. The sentences were annotated based on real-world knowledge, e.g. a sentence like ``Muhammed is like Hitler'' would be annotated as a negative sentiment towards Muhammed. Further, if a commenter presented a negative fact about the keyword, the sentence would be denoted as negative.

@@

Finally the material was divided into two parts where the first half of the debate articles with subsequent discussions make up the training set and the rest constitutes a held-out test set. In the development of the important dependency paths, only the training set was used. After the division, the training and test set consisted of a total of 382 and 308 sentences, respectively. Table \ref{tab:1} summarizes the sentiments in the sentences.
\begin{table}
  \centering
  \begin{tabular}{lcccc}
             & Negative & Neutral & Positive \\\hline
    Training & 174 (46\%) & 162 (42\%) & 46 (12\%)\\
    Test     & 102 (33\%)& 182 (59\%) & 24 (8\%)
  \end{tabular}
  \label{tab:1}
  \caption{Manual annotation of training and test set.}
\end{table}

\subsection{Corpus postprocessing}
The sentiment corpus was PoS-tagged and parsed using the
Bohnet\&Nivre-parser \cite{Boh:Niv:12}. This parser is a
transition-based dependency parser with joint tagger that implements
global learning and a beam search for non-projective labeled
dependency parsing.  This latter parser has recently outperformed
pipeline systems (such as the Malt and MST parsers) both in terms of
tagging and parsing accuracy for typologically diverse languages such
as Chinese, English, and German. It has been reported to obtain a
labeled accuracy of 87.7 for Norwegian \cite{Sol14}.  The parser is
trained on the Norwegian Dependency Treebank (NDT). The NDT is a
treebank created at the National Library of Norway in the period
2011-2013, manually annotated with part-of-speech tags, morphological
features, syntactic functions and dependency graphs \cite{Sol14,
 Sol13}. It consists of approximately 600 000 tokens, equally
distributed between Norwegian Bokm{\aa}l and Nynorsk, the two Norwegian
written standards. Only the Bokm{\aa}l subcorpus has been used here. A
large proportion of the NDT is newspaper text, there are also
parliament transcripts, government reports and texts with a more
colloquial style from blogs. Detailed annotation guidelines in English
will be made available in April 2014 \cite{Kinn2013}.

@@

Several Norwegian sentiment shifters were considered but only the basic shifter 'not' improved the sentiment classification and therefore only this word were used in the method.

\section{Experiments}

In this study we compared four different methods based on the general algorithm in \eqref{eq:1}.
\begin{itemize}
\item We use the \textbf{imp}-function presented in \eqref{eq:2}. We denote this method WD (word distance).
\item For this method and the two below we use the \textbf{imp}-function in \eqref{eq:3}. Further we set $\alpha = 0$ which means that we do not use the important dependency paths. We denote this method A0 ($\alpha = 0$).
\item We set $\alpha = 1$ and for all dependency paths we set $W_{\mathbf{dep}} = 2/3$. We denote this method CW (constant weights).
\item We set $\alpha = 1$ and for $W_{\mathbf{dep}}$ we use the weights presented in Table ??????. We denote this method OD (optimal use of dependency paths)
\end{itemize}
For each method we used the training set to manually tune the parameters $d_p, d_n, t_p$ and $t_n$ of the method. The parameters were tuned to optimize the number of correct classifications.

Table \ref{tab:2} shows the optimal parameter values of $d_p, d_n, t_p$ and $t_n$ tuned using the training set, and classification performance for the different methods on the test set using the parameter values tuned from the training set. The p-values are computed using the regression model presented in Section \ref{sec:sa}.
\begin{table}
  \caption{The second to the fifth column show the optimal values of the parameters of the model tuned using the training set. The sixth column show the number of correct classifications and the last column shows p-values testing whether the method performs better than WD.}
  \centering
  \begin{tabular}{lcccccl}
       & $d_p$ & $d_n$ & $t_p$ & $t_n$ & Correct   & p-val \\ \hline
    WD &    2  &   0  & 0.7  &  0.0  & 145 (47\%) & \\
    A0 &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.023\\
    CW &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.024\\
    OD &    2  &   0  & 2.0  &  0.3  & 162 (53\%) & 0.016
  \end{tabular}
  \label{tab:2}
\end{table}
We see that the sentiment shifter 'not' only have a positive effect on the classification performance when it is in front of the sentiment word. We see that using dependency distances (method A0) the classification results is significantly improved compared to using word distances in the sentence (method WD) (p-value = 0.023). Also classification based on important dependency paths (method OD) performs significantly better than WD. We also see that OD performs better than A0 (162 correct compared to 161), but this improvement is not statistically significant.

@@

Classifying sentiment in political discussions is hard because of the frequent use of irony, sarcasm and citations. In this paper we have compared the use of word distance between keyword and sentiment word against metrics related to parsed sentence information. Our results show that using dependency tree distances or important dependency paths, improves the classification performance compared to using word distance.

>>>>

%keywords: \textit{}

@@

In this paper we focus on classifying the sentiment towards
religious/political topics, say the Quran, in Norwegian political
discussion. We use a lexicon-based approach where we classify the
sentiment of a sentence based on the polarity of sentiment words in
relation to a set of target words in the sentence. We expect that
statistically the importance of a sentiment word towards the target
word is related to the number of words between the sentiment and
target word as suggested by Ding et al. \shortcite{Ding08}.
Information about the syntactic environment of certain words or
phrases has in previous work also been shown to be useful for the task
of sentiment classification \cite{Wil:Wie:Hof:09,Jiang11}. In this
work we therefore compare the results obtained using a token-based
distance measure with a novel syntax-based distance measure obtained
using dependency graphs and further augmented with linguistically
motivated syntactic patterns expressed as dependency paths. In order
to evaluate the proposed methods, we furthermore present a freely
available corpus of Norwegian political discussion related to religion
and immigration, which has been manually annotated for the sentiment
expressed towards a set of target words, as well as a manually
translated sentiment lexicon.

%% In this paper we focus on classifying the opinion toward religious/political topics, say the Quran, in political discussion by using the lexical-based approach. One intuitive approach is to find both the target word (e.g. Quran) and the words with sentiment in the sentence and classify the sentiment of the sentence based on the polarity of these sentiment words. We expect that statistically the importance of a sentiment word towards the target word is related to the number of words between the sentiment and key word as suggested by Ding et al. \shortcite{Ding08}. Two other approaches is to automatically parse the material and either use the distance between key and sentiment word in the parse tree or develop grammatical dependence paths, see e.g. Jiang et al. \shortcite{Jiang11}. The aim of this paper is to compare the performance of a word distance method \cite{Ding08} with novel methods based on distance in parse tree and grammatical dependence paths to classify opinions in political discussions.

@@

In this section we present two methods to classify sentences as either positive, neutral or negative towards a target word. Both methods follow the same general algorithm presented below which is inspired by Ding et al. \shortcite{Ding08} and is based on a list of sentiment words each associated with a sentiment score representing the polarity and strength of the sentiment word (sentiment lexicon). Both target words, sentiment words and sentiment shifters can in general appear several times in a sentence. Sentiment shifters are words that potentially shift the sentiment of a sentence from positive to negative or negative to positive. E.g. ``not happy'' have the opposite polarity than just ``happy''. Let $kw_i, i \in \{1,2,\ldots,I\}$ represent appearance number $i$ of the target word in the sentence. Further let $sw_{j}, j \in \{1,2,\ldots,J\}$ be appearance number $j$ of a sentiment word in the sentence. Finally let $\textbf{ss} = (ss_1, ss_2, \ldots, ss_K)$ repressent the sentiment shifters in the sentence. We compute a sentiment score, $S$, for the sentence as follows
\begin{equation}
  \label{eq:1}
  S = \frac{1}{I}\sum_{i=1}^{I} \sum_{j=1}^{J} \mathbf{imp}(kw_i, sw_{j})\mathbf{shift}(sw_{j}, \mathbf{ss})    
\end{equation}
where the function \textbf{imp} computes the importance of the sentiment word $sw_{j}$ on the target word appearance $kw_i$. This will be computed in different ways as described below. Further, the function $\mathbf{shift}(sw_{j}, \mathbf{ss})$ computes whether the sentiment of $sw_{j}$ should be shifted based on all the sentiment shifters in the sentence. It returns $-1$ (sentiment shift) if some of the sentiment shifters are within $d_{p}$ words in front or $d_{n}$ words behind $sw_{j}$, respectively. Else the function, returns $1$ (no sentiment shift). We classify the sentiment towards the target word to be positive, neutral or negative if $S >= t_p$,  $t_p > S > t_n$ and  $S <= t_n$, respectively. The parameters $d_p, d_n, t_p$ and $t_n$ is tuned using a training set, as described in section \ref{sec:exp} below.

@@

\paragraph{Dependency distance} One way of expressing the syntactic environment of a target word with respect to a sentiment word is to determine its distance in the dependency graph. We therefore define a distance function $\mathbf{depdist}(kw_i, sw_{j})$ which returns the number of nodes in the shortest dependency path from the target word to the sentiment word in the dependency graph. The shortest path is determined using Dijkstra's shortest path algorithm \cite{Dij:59}.

\paragraph{Dependency paths}
A second way of determining the importance of a sentiment word towards a target based on syntactically parsed texts, is to establish a list of grammatical dependency paths between words, and test whether such paths exist between the targets and sentiment words \cite{Jiang11}. The assumption would be that two words most likely are semantically related to each other if there is a meaningful grammatical relation between them. Furthermore, it is reasonable to expect than some paths are stronger indicators of the overall sentiment of the sentence than others. To test this method, we have manually created a list of 42 grammatical dependency paths, divided into four groups, and given them a score between 1-3. The higher the score is, the better indicator of sentiment the path is assumed to be. In the following paragraphs, we will briefly present the groups of paths and the maximum score we have assigned in each group. The paths are represented in the following format: postag-target:postag-sentiment word{\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn({\textunderscore}{\textunderscore}DEPREL{\textunderscore}up/dn etc.). \emph{Up} and \emph{dn} indicate the direction of the traversed arc in the graph.

@@

A second group we have considered, contains paths from subject targets to sentiment words embedded within the predicate, such as from the subject to the nominal direct object of a verb, subst:subst{\textunderscore}{\textunderscore}SUBJ{\textunderscore}up{\textunderscore}{\textunderscore}DOBJ{\textunderscore}dn. Paths from subjects into different kinds of adverbials are also a part of this group. We consider paths from subjects to objects to be good indicators of sentiment and assign them the highest score, 3, based on the same arguments as for subject-predicate paths. Paths into adverbials get lower values, as adverbials often are less semantically connected to the predicate than objects.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{|l|l|l|}
\hline \bf Path group & \bf Number & \bf Score range \\ \hline
Subj. to pred. & 9 & 3 \\
Subj. to pred.-internal & 13 & 3-1 \\
Pred.-internal & 6 & 1 \\
NP-internal & 14 & 2-1 \\
\hline
\end{tabular}
\end{center}
\caption{\label{deppath} Grouping of dependency paths with the number of paths and score range for each group.}
\label{tab:paths}
\end{table}

\paragraph{Modified scoring function}
Let $\mathcal{D}$ denote the set of all important dependency paths. The function $\mathbf{gram}(kw_i, sw_{j})$ returns the dependency path, and if $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$, then the function $W_{\mathbf{dep}}(kw_i, sw_{j}) \in [0,1]$, returns the importance of the path. Further let $\mathbf{depdist}(kw_i, sw_{j})$ return the dependency distance, as described above. The \textbf{imp} function is computed as follows. If $\mathbf{gram}(kw_i, sw_{j}) \in \mathcal{D}$ we use
\begin{align}
  \begin{split}
    \label{eq:3}
  \mathbf{imp}&(kw_i, sw_{j}) = \\
  &\alpha \cdot \mathbf{sentsc}(sw_{j}) W_{\mathbf{dep}}(kw_i, sw_{j}) \\
  +&(1 - \alpha) \cdot \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{depdist}(kw_i, sw_{j})}    
  \end{split}
\end{align}
where $\alpha \in [0,1]$ is a parameter that weights the score from the important dependency path and the tree distance and can be tuned using a training set. If $\mathbf{gram}(kw_i, sw_{j}) \not\in \mathcal{D}$ we simply use 
\begin{equation}
  \label{eq:4}
  \mathbf{imp}(kw_i, sw_{j}) = \dfrac{\mathbf{sentsc}(sw_{j})}{\mathbf{depdist}(kw_i, sw_{j})}    
\end{equation}
Note that when $\alpha = 0$, \eqref{eq:3} reduces to \eqref{eq:4}.

@@

We did not find any suitable annotated text material related to political discussions in norwegian and therefore created our own. We manually selected 46 debate articles from the Norwegian online newspapers \textit{NRK Ytring, Dagbladet, Aftenposten, VG and Bergens Tidene}. To each debate article there were attached a discussion thread where readers could express their opinions and feelings towards the content of the debate article. All the text from the debate articles and the subsequent discussions were collected using text scraping \cite{Hammer13}. The debate articles were related to religion and immigration and we wanted to classify the sentiment towards all forms of the following target words: \textit{islam, muslim, quran, allah, muhammed, imam and mosque}. These are topics that typically create a lot of active discussions and disagreements.

We automatically divided the material into sentences and all sentences containing at least one target word and one sentiment word were kept for further analysis. If a sentence contained more than one target word, e.g. both Islam and Quran, the sentence was repeated one time for each target word in the final text material. We could then classify the sentiment towards each of the target words in the sentence consecutively. To assure that we do not underestimate the uncertainty in the statistical analysis, we see each repetition of the sentence as the same sentence with respect to the sentence random effect in the regression model in Section \ref{sec:sa}

Each sentence was manually annotated as to whether the commenter was positive, negative or neutral towards the target word in the sentence. Each sentence was evaluated individually. The sentences were annotated based on real-world knowledge, e.g. a sentence like ``Muhammed is like Hitler'' would be annotated as a negative sentiment towards Muhammed. Further, if a commenter presented a negative fact about the target word, the sentence would be denoted as negative.

@@

Finally the material was divided into two parts where the first half of the debate articles with subsequent discussions make up the training set and the rest constitutes a held-out test set. In the manual development of the important dependency paths, only the training set was used. After the division, the training and test set consisted of a total of 382 and 308 sentences, respectively. Table \ref{tab:1} summarizes the annotation found in the corpus.
\begin{table}
  \centering
  \begin{tabular}{lcccc}
             & Negative & Neutral & Positive \\\hline
    Training & 174 (46\%) & 162 (42\%) & 46 (12\%)\\
    Test     & 102 (33\%)& 182 (59\%) & 24 (8\%)
  \end{tabular}
  \label{tab:1}
  \caption{Manual annotation of training and test set.}
\end{table}

\subsection{Corpus postprocessing}
The sentiment corpus was PoS-tagged and parsed using the
Bohnet\&Nivre-parser \cite{Boh:Niv:12}. This parser is a
transition-based dependency parser with joint tagger that implements
global learning and a beam search for non-projective labeled
dependency parsing.  This latter parser has recently outperformed
pipeline systems (such as the Malt and MST parsers) both in terms of
tagging and parsing accuracy for typologically diverse languages such
as Chinese, English, and German. It has been reported to obtain a
labeled accuracy of 87.7 for Norwegian \cite{Sol14}.  The parser is
trained on the Norwegian Dependency Treebank (NDT). The NDT is a
treebank created at the National Library of Norway in the period
2011-2013, manually annotated with part-of-speech tags, morphological
features, syntactic functions and dependency graphs \cite{Sol14,Sol13}. It consists of approximately 600 000 tokens, equally
distributed between Norwegian Bokm{\aa}l and Nynorsk, the two Norwegian
written standards. Only the Bokm{\aa}l subcorpus has been used here. A
large proportion of the NDT is newspaper text, there are also
parliament transcripts, government reports and texts with a more
colloquial style from blogs. Detailed annotation guidelines in English
will be made available in April 2014 \cite{Kinn2013}.

@@

Several Norwegian sentiment shifters were considered but only the basic shifter 'not' improved the sentiment classification and therefore only this word was used in the method.

\section{Experiments}
\label{sec:exp}
In this study we compare four different methods based on the general algorithm in \eqref{eq:1}.
\begin{itemize}
\item We use the \textbf{imp}-function presented in \eqref{eq:2}. We denote this method WD (word distance).
\item For this method and the two below we use the \textbf{imp}-function in \eqref{eq:3}. Further we set $\alpha = 0$ which means that we do not use the important dependency paths. We denote this method A0 ($\alpha = 0$).
\item We set $\alpha = 1$ and for all dependency paths we set $W_{\mathbf{dep}} = 2/3$. We denote this method CW (constant weights).
\item We set $\alpha = 1$ and for $W_{\mathbf{dep}}$ we use the weights presented in Table \ref{tab:paths}. We denote this method OD (optimal use of dependency paths)
\end{itemize}
For each method we used the training set to manually tune the parameters $d_p, d_n, t_p$ and $t_n$ of the method. The parameters were tuned to optimize the number of correct classifications.

Table \ref{tab:2} shows the optimal parameter values of $d_p, d_n, t_p$ and $t_n$ tuned using the training set, and classification performance for the different methods on the test set using the parameter values tuned from the training set. The p-values are computed using the regression model presented in Section \ref{sec:sa}.
\begin{table}
  \centering
  \begin{tabular}{lcccccl}
       & $d_p$ & $d_n$ & $t_p$ & $t_n$ & Correct   & p-val \\ \hline
    WD &    2  &   0  & 0.7  &  0.0  & 145 (47\%) & \\
    A0 &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.023\\
    CW &    2  &   0  & 2.0  &  0.3  & 161 (52\%) & 0.024\\
    OD &    2  &   0  & 2.0  &  0.3  & 162 (53\%) & 0.016
  \end{tabular}
  \caption{The second to the fifth column show the optimal values of the parameters of the model tuned using the training set. The sixth column show the number of correct classifications and the last column shows p-values testing whether the method performs better than WD.}
  \label{tab:2}
\end{table}
We see that the sentiment shifter 'not' only has a positive effect on the classification performance when it is in front of the sentiment word. We see that using dependency distances (method A0) the classification results are significantly improved compared to using word distances in the sentence (method WD) (p-value = 0.023). Also classification based on important dependency paths (method OD) performs significantly better than WD. We also see that OD performs better than A0 (162 correct compared to 161), but this improvement is not statistically significant.

@@

Classifying sentiment in political discussions is hard because of the frequent use of irony, sarcasm and citations. In this paper we have compared the use of word distance between target word and sentiment word against metrics incorporating syntactic information. Our results show that using dependency tree distances or important dependency paths, improves the classification performance compared to using word distance.
